---
title: "A3: SPL Library Checkouts"
output: html_document
---

## Project title

Trend of the checkouts number over time 

### Authors

- Qin Deng, qdeng2@uw.edu

### Date

2023 2/24
### Introduction

Libraries have always played a vital role in society by providing access to information and resources for individuals to improve their education, skills, and knowledge. With the increasing digitization of books and other resources, libraries have been adapting to new trends in borrowing and lending. The Monthly Item Borrowing Trends data from the library provides valuable insights into how patrons interact with the library's collection. This data includes information on the monthly borrowing trends of various items, including print books, audio books, eBooks, and more, from 2017 to 2023. Analyzing this data can help library administrators and staff better understand patrons' borrowing habits, identify popular and less popular items, and plan for future acquisitions and services. In this project, I will use R programming language to analyze this data, visualize trends over time, and gain insights into borrowing patterns. And I calculate the data of LW_avg_checkout, JE_avg_checkout, JE_max_year, LM_max_year and EB_max_year.

I will analyze the checkout times in our group project. I am going to create some data visualization to help people compare and contrast data. Here are the questions that my project can answer.
- What is the average checkout number for Little Women?
- What is the average checkout number for Jane Eyre? 
- What is the year has the highest checkout number for Jane Eyre?
- What is the year has the highest checkout number for Little Women?
- What is the year has the highest checkout number for e-books?

### Summary Information

```{r, echo = FALSE, message= FALSE, warning = FALSE, code = readLines("summary.R")}
source("summary.R") 
```

The first data I studied is the average checkout number for Little Women, which is the `r LW_avg_checkout` times. In the meantime, I also calculate the average number of Jane Eyre, which is the `r JE_avg_checkout`times. The reason I made them the subject of my research is that they share the same subject matter about the Feminist. And also I think the reason that Jane Eyre has a larger average checkout number is because the Jane Eyre movie aired and made people want to read the original book. Therefore I think it is good for us to compare them together. The third and forth values are about the year has the highest checkout number for Jane Eyre and Little Women, which are `r JE_max_year` and `r LM_max_year`. I believe these two statistics prove that more and more people are taking Woman's Rights seriously and want to read books about them to learn more about them. 
The last value is the year has the highest checkout number for e-books, which is `r EB_max_year`.  I believe this trend can be attributed, in part, to the COVID-19 pandemic, which made it difficult for people to physically visit libraries. E-books offer a convenient and safe alternative that allows people to read from the comfort of their own homes, contributing to their growing popularity.

### The Dataset

The Seattle Public Library has provided valuable data on the physical and electronic book checkouts. This data was collected through the library's circulation system, which automatically records the number of checkouts for each object. The primary purpose of the data is to keep track of the total amount and dates of checked books and to inform collection development decisions and initiatives aimed at promoting literacy and information access. The dataset contains 42.2M rows and 12 columns, providing extensive information on various categories of books and their circulation dates.

However, working with this data requires careful consideration of ethical concerns. The data may contain information about the checkout histories of specific customers, which raises privacy concerns. To protect individuals' privacy, the data must be anonymized before analysis. Moreover, there is a risk of drawing inferences or perpetuating stereotypes based on the materials most commonly borrowed, which should be avoided.

Despite its significance, the data also has some limitations. It only includes information on checkouts from the Seattle Library system and does not cover other libraries or reading material sources. This means that the findings from this data may not be representative of the wider population's reading habits.

### First Trends Over Time Chart


```{r, echo = FALSE, message= FALSE, warning = FALSE, code = readLines("Chart_1.R")}
source("Chart_1.R")
line_chart1
```

According to the graph, we can see "Fiction, Literature, Romance" has a larger amount of checkouts than others. Therefore, we can conclude that people prefer to read articles of Fiction, Literature, and Romance. Studying the trend of checkout items' subjects over time is to gain insights into the changing interests and preferences of library patrons. In order to make decisions about collection development and resource allocation, researchers can identify patterns and trends in readers' reading habits by examining the kinds of materials that people borrow from libraries. Researchers can learn more about the reading preferences of library users by looking at the categories that see more checkouts, like Fiction, Literature, and Romantic. For instance, it might imply that readers favor certain genres over others and are more apt to read for enjoyment.

### Second Trends Over Time Chart

```{r, echo = FALSE, message= FALSE, warning = FALSE, code = readLines("Chart_2.R")}
source("Chart_2.R")
line_chart2
```
Based on the graph, it is evident that there has been a surge in the popularity of various types of materials from 2020 to 2022. One probable reason for this trend could be the impact of COVID-19, which has restricted people's outdoor activities, leading to an increased demand for home-based activities such as reading books, listening to audio books, and watching videos. Notably, physical books have witnessed a significant decline in checkouts, moving from the highest in 2018 to the second-lowest by 2022. Consequently, this data could aid libraries in tailoring their services to cater to the evolving trends in different material types. For instance, they can increase their collection of eBooks to meet the changing demands of their patrons. Furthermore, as consumer interests continue to shift away from videodiscs and books, libraries could consider reducing the purchase of such items and the number of staff assigned to serve them, thus enhancing cost-effectiveness.

### Your Choice


```{r, echo = FALSE, message= FALSE, warning = FALSE, code = readLines("Chart_3.R")}
source("Chart_3.R")
line_chart3
```

I included the chart to visually represent the trend of digital and physical items in the library over the years. The chart helps to showcase the changes in the popularity of each type of material and provides a clear comparison between the two. It helps to illustrate the significant decline in the use of physical items and the corresponding increase in the use of digital items. The chart reveals that in 2018, physical items were more popular than digital items in the library, with more than twice as many physical items checked out compared to digital items. However, from 2020 onwards, there has been a sharp decline in the use of physical items, with the number dropping to only 1/5 of the 2018 figure by 2022. In contrast, the use of digital items has been on a steady upward trend, with the number doubling from 2018 to 2022. The chart also highlights that the popularity of physical items has been decreasing steadily, while the popularity of digital items has been increasing at a faster rate. This trend suggests a shift in reading habits and library usage towards more convenient and accessible digital materials. Analyzing this trend can assist libraries in understanding the shift in UsageClass patterns, enabling them to reduce the supply of physical items and increase the supply of digital items. This strategic move can potentially result in increased profitability for libraries by catering to the changing needs and preferences of their patrons.

install.packages('rsconnect')
library(rsconnect)
